---
title: "Exploring Gun Deaths in America"
author: "Cheng Yee Lim"
output: github_document
---

### Get the data

```{r, results='hide'}
library(tidyverse)    # load tidyverse packages, including ggplot2
library(rcfss)
library(knitr)
library(stringr)
library(scales)
library(lvplot)

# load the data
data("gun_deaths")
gun_deaths
```

## In what month do the most gun deaths occur?
Most gun deaths occur in July. 
```{r}
count(gun_deaths, month) %>%
  kable(caption = "Gun Deaths in the United States (2012-2014) by Month", 
        col.names = c("Month", "Number of Deaths"))
```

### With a bar chart 
```{r}
ggplot(data = gun_deaths) + 
  geom_bar(aes(month, fill = factor(month))) + 
  ggtitle("Gun Deaths in the United States (2012-2014) by Month") + 
  xlab("Month") + 
  ylab("Number of Gun Deaths") + 
  scale_fill_discrete(guide=FALSE) +
  scale_x_continuous(
    breaks = 
      round(seq(min(gun_deaths$month), max(gun_deaths$month), by = 1.0),1), 
    labels = 
      c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
               "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) 
```

### With a line chart 
```{r}
month.gun <- gun_deaths %>%
  group_by(month) %>%
  summarize(count = n())

ggplot(data=month.gun) + 
  geom_line(aes(month, count)) +
  ggtitle("Gun Deaths in the United States (2012-2014) by Month") + 
  xlab("Month") + 
  ylab("Number of Gun Deaths") + 
  ylim(0, 9000) + 
  scale_x_continuous(
    breaks = 
      round(seq(min(gun_deaths$month), max(gun_deaths$month), by = 1.0),1), 
    labels = 
      c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
        "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
```

## What is the most common intent in gun deaths? Do most people killed by guns die in suicides, homicides, or accidental shootings?
The most common intent in gun deaths is suicide. 
```{r}
count(gun_deaths, intent) %>%
  kable(caption = "Gun Deaths in the United States (2012-2014) by Intent", 
        col.names = c("Intent of Shooting", "Number of Deaths"))
```

### With a bar graph 
```{r}
intent.gun <- filter(gun_deaths, intent!= "NA")
ggplot(data = intent.gun) + 
  geom_bar(aes(intent, fill = factor(intent))) + 
  ggtitle("Gun Deaths in the United States (2012-2014) by Intent of Death") + 
  xlab("Intent of Death") + 
  ylab("Number of Gun Deaths") + 
  scale_fill_discrete(guide=FALSE) +
  scale_x_discrete(
    limits=c("Suicide","Homicide","Accidental", "Undetermined"))
```

## What is the average age of females killed by guns?
The average age of females killed by guns is 43.74. 
```{r}
female.age.guns <- gun_deaths %>%
  group_by(sex) %>%
  filter(age > 0) %>%
  summarize(female.age = mean(age, na.rm = TRUE))

kable(female.age.guns, 
      caption = "Average Age of Gun Death Victims (2012-2014) by Sex", 
      col.names = c("Sex", "Average Age"), digits = 2)
```

### Using a bar graph 
```{r}
ggplot(data = female.age.guns) + 
  geom_bar(aes(sex, female.age), stat = "identity") + 
  ggtitle("Average Age of Gun Deaths Victims (2012-2014) by Sex") +
  xlab("Sex of Victim") + 
  ylab("Age of Victim") + 
  scale_x_discrete(breaks =c("F", "M"), labels = c("Female", "Male"))
```

### Using a boxplot 
```{r}
ggplot(data=gun_deaths) + 
  geom_boxplot(aes(sex, age)) + 
  ggtitle("Average Age of Gun Deaths Victims (2012-2014) by Sex") + 
  xlab("Sex of Victim") + 
  ylab("Age of Victim") +
  scale_x_discrete(breaks =c("F", "M"), labels = c("Female", "Male"))
```

### Using a violin plot 
```{r}
ggplot(data=gun_deaths) + 
  geom_violin(aes(sex, age)) + 
  ggtitle("Average Age of Victims (2012-2014) by Sex") + 
  xlab("Sex") + 
  ylab("Age of Victim") +
  scale_x_discrete(breaks =c("F", "M"), labels = c("Female", "Male"))
```

### Using a letter-value boxplot 
``` {r}
ggplot(data=gun_deaths) + 
  geom_lv(aes(sex, age)) + 
  ggtitle("Average Age of Victims (2012-2014) by Sex") + 
  xlab("Sex") + 
  ylab("Age of Victim") +
  scale_x_discrete(breaks =c("F", "M"), labels = c("Female", "Male"))
```

## How many white males with at least a high school education were killed by guns in 2012?
There were 15199 white males with at least a high school education killed by guns in 2012.

```{r}
white.highschool <- gun_deaths %>% 
  filter(education!="NA") %>%
  filter(sex == "M") %>%
  filter(race == "White") %>%
  filter(year == 2012) %>%
  filter(education != "Less than HS") %>% 
  summarize(count = n())
```

## Which season of the year has the most gun deaths?

Assuming that:

* Winter = January-March
* Spring = April-June
* Summer = July-September
* Fall = October-December

Summer season has the most gun deaths. 

```{r}
gun_deaths$season <- cut(gun_deaths$month,
                         breaks = c(-Inf, 3, 6, 9, Inf),
                         labels = c("Winter","Spring","Summer", "Fall"))

season.gun <- gun_deaths %>%
  group_by(season) %>%
  summarize(count = n())

kable(season.gun, 
      caption = "Gun Deaths in the United States (2012-2014) by Season",
      col.names = c("Season", "Number of Deaths"), digits = 2)

ggplot(data = gun_deaths) + 
  geom_bar(aes(season, fill = season)) + 
  ggtitle("Gun Deaths in the United States (2012-2014) by Season") + 
  xlab("Season") + 
  ylab("Number of Gun Deaths") + 
  scale_fill_discrete(guide=FALSE)

```

## What is the relationship between race and intent?

Whites who are killed by guns are more likely to die because of suicide, whereas blacks and hispanics are more likely to die because of homicide. 

```{r}
gun_deaths$newrace <- str_wrap(gun_deaths$race, width = 10)

intent.gun <- gun_deaths %>%
  group_by(race, intent) %>%
  summarize(count= n())

kable(intent.gun, col.names = c("Race of Victim", "Intent of Shooting", "Number of Deaths"))
```

### Using a horizontal bar graph 
```{r}
ggplot(data = gun_deaths) + 
  geom_bar(aes(x =newrace, fill=intent), position = "fill") + 
  ggtitle("Gun Deaths in the United States (2012-2014) by Race and Intent") + 
  xlab("Race of Victim") + 
  ylab("Percent of Gun Death Victims") + 
  labs(fill = "Intent") +
  theme(legend.position = "bottom") +
  coord_flip() + 
  ylim(0, 70000) + 
  scale_y_continuous(labels = percent_format())
```

## Are police-involved gun deaths significantly different from other gun deaths? Assess the relationship between police involvement and age, police involvement and race, and the intersection of all three variables.
The age of police-involved gun death victims is often younger than other victims.

```{r, results='hide'}
cut(gun_deaths$police, breaks = c(-.99,0,1))
gun_deaths$police <- factor(gun_deaths$police)
```

### Investigating police involvement and average age of victims with a boxplot
```{r}
ggplot(data=gun_deaths) + 
  geom_boxplot(aes(x = police, y= age, fill='cyl')) + 
  ggtitle("Average Age of Victims (2012-2014) by Police Involvement") +
  ylab("Age of Victim") +
  xlab("") + 
  scale_x_discrete(
    breaks = c(0,1), 
    labels = c("No police involved","Police Involved")) + 
  scale_fill_discrete(guide=FALSE)
```

### Investigating percentage of gun death victims by race and policement involvement with a bar graph 
There are significantly higher percentage of Hispanic gun death victims, who died due to an officer-involved shooting, when compared to percentage of Hispanic victims who died without police involvement. Conversely, the percentage of victims who were white in a police-involved gun death were smaller than percentage of White victims who died without police involvement.

```{r}
gun_deaths$race <- str_wrap(gun_deaths$race, width = 15)
gun_deaths$race <- factor(gun_deaths$race)

ggplot(data = gun_deaths) + 
  geom_bar(aes(x=police, fill=race), position = "fill") + 
  ggtitle("Percentage of Gun Death Victims by Race and Police Involvement") +
  labs(fill = "Race") + 
  ylab("Percent of Gun Death Victims") + 
  xlab("") +
  scale_y_continuous(labels = percent_format()) + 
  scale_x_discrete(breaks = c(0,1), 
                     labels = c("No police involved","Police Involved"))
```

### Relationship between police involvement, race and age
With the exception of the Black gun death victims, the average age of victims who died as a result of an officer-involved shooting was younger than victims who died without police involvement. 
```{r, results ='hide'}
gun_deaths$rac[gun_deaths$race=="Asian/Pacific Islander"] <- 1
gun_deaths$rac[gun_deaths$race=="Black"] <- 2
gun_deaths$rac[gun_deaths$race=="Hispanic"] <- 3
gun_deaths$rac[gun_deaths$race=="Native American/Native Alaskan"] <- 4
gun_deaths$rac[gun_deaths$race=="White"] <- 5
cut(gun_deaths$rac, breaks = c(0,1,2,3,4,5))
```

```{r}
ggplot(data = gun_deaths) + 
  geom_boxplot(aes(x=race, y=age, fill=police)) + 
  ggtitle("Relationship between Age, Race and Police Involvement") +
  xlab("Race of Victim") + 
  ylab("Age of Victim") +
  theme(legend.position = "bottom") +
  scale_fill_discrete(name = "Shooting Type",
                      breaks = c("0", "1"),
                      labels = c("No Police Involved", "Police Involved")) +
  coord_flip()

ggplot(data = gun_deaths) + 
  geom_boxplot(
    aes(x=police, y=age, fill=race)) + 
  ggtitle("Relationship between Age, Race and Police Involvement") +
  xlab("Shooting Type") + 
  ylab("Age of Victim") +
  labs(fill = "Race of Victim") +
  scale_x_discrete(breaks = c(0,1), 
                     labels = c("No police involved","Police Involved")) +
  theme(legend.position = "bottom") 
```

## Session info

```{r}
# don't modify this chunk
devtools::session_info()
```


